---
title: "class7 lab (Hands on with Principal Component Analysis (PCA))"
author: "Josef Urrete"
format: pdf
---

# 1. PCA of UK Food data

```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url)

```

## Q1. How mnay rows and columns are in dataframe x? What R functions could you use to answer these questions?

```{r}
#use nrows and ncol or dim()
dim(x)
```

```{r}
#to preview x
##View(x)
```

```{r}
#how minus indexing works
rownames(x) <- x[,1]
x <- x[,-1]
head(x)
```

```{r}
x <- read.csv(url, row.names = 1)
head(x)
```

## Q2: Which approach to solving the "row-names problem' mentioned above do you prefer and why? Is one approach more robust than another under certain circumstances?

Making basic visualizations

```{r}
cols <- rainbow(nrow(x))
barplot(as.matrix(x), col=cols)
```

## Q3: Chainging optional argument in barplot

```{r}
barplot(as.matrix(x), col=cols, beside = T)
```

```{r}
pairs(x, col =rainbow(nrow(x)), pch = 16)
```

Applying PCA

```{r}
pca <- prcomp(t(x))
attributes(pca)
```

## Q7: complete the code below to generate a plot of PC1 vs PC2. Second line adds text labels over the data points.

using pca\$x

```{r}
plot(pca$x[,1], pca$x[,2])
text(pca$x[,1], pca$x[,2], colnames(x))
```

## Q8: Customize plot so colors of country names match colors in original map and table

```{r}
country_cols <- c("orange", "red", "blue", "green")
plot(pca$x[,1],pca$x[,2])
text(pca$x[,1], pca$x[,2], colnames(x), col=country_cols)
```

PC1

```{r}
par(mar=c(10,3,0.35,0))
barplot(pca$rotation[,1], las=2)
```

## Q9: Generate similar loadings plot for PC2. What two food groups feature prominantly and what does PC2 mainly tell us about?

```{r}
par(mar=c(10, 3, 0.35, 0))
barplot(pca$rotation [,2], las=2)

```

```{r}
biplot(pca)
```

```{r}
url2 <- "https://tinyurl.com/expression-CSV"
rna.data <- read.csv(url2, row.names=1)
head(rna.data)
```

## Q10: how many genes and samples are in this ddata set?

```{r}
dim(rna.data)
```

100 genes and 10 samples

```{r}
pca <- prcomp(t(rna.data))
summary(pca)
```

```{r}
plot(pca$x[,1], pca$x[,2], xlab = "PC1", ylab ="PC2")
```

```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2")
text(pca$x[,1], pca$x[,2], colnames(rna.data))
```

```{r}
sort(pca$rotation[,1])
```
